---
format:
  revealjs:
    theme: night
    embed-resources: true
    slide-number: true
    navigation-mode: linear
    controls: true
    auto-stretch: false 
    code-overflow: wrap
    incremental: true
    css: styles.css
    fontsize: "25pt"
---

# Code {background-color="#E77500"}

# Topics

1.  Code chunks
2.  Inline code

## What are code chunks?

-   Specified place for code
-   Executable
-   Weaved between text
-   May print output or create new variables, for example

<br>

. . .

```{r}
#| echo: true
1993 + 2023

number=1993+2023
number
```

## Anatomy of a code chunk

::::: columns
::: {.column width="50%"}
1.  On each end of the code chunk: 3 *backticks*: ```` ``` ```` \*\*
2.  Curly brackets: *engine*: `{r}` or `{python}` \*\*
3.  Top of chunk: Code chunk options, with *hashpipe*: `#|`
:::

::: {.column width="50%"}
```` md
```{{r}}
#| label: cars-r
#| echo: false
mtcars %>% 
  distinct(cyl)
```
````

<br>

```` md
```{{python}}
#| label: car-py
#| echo: false
distinct_cyl = mtcars.drop_duplicates(subset=['cyl'])
```
````
:::
:::::

\*\* *not needed in a code cell in a Jupyter notebook!*

## Code chunk options {.smaller}

Useful [code chunk options](https://quarto.org/docs/computations/execution-options.html)

| Option | Description |
|----|----|
| `eval` | `Evaluate the code chunk.` |
| `echo` | `Include the source code in output` |
| `output` | \`Include code output results (`true`, `false`, or `asis`)\` |
| `warning` | `Include warnings in the output.`\` |
| `error` | `Include errors in the output (continues execution if error present).` |
| `include` | `Catch all for preventing any output (code or results) from being included.` |
| `label` | `Please make sure to label your code chunks! It helps with debugging.` |
| `fig-cap` | `caption your figure` |
| `fig-width` | `width of the figure` |
| `fig-height` | `height of the figure` |

## Code Chunk Options

Options for *all* code chunks: include in `yaml` (top of the document).

. . .

For example...

``` yaml
---
format: html
execute:
  echo: false
---
```

... hides all code in the output!

## Inserting Code Chunks

Option 1: Source editor: Type it out

. . .

Option 2: Visual editor: `Insert` \> `Code cell`

. . .

Option 3: Shortcuts:

-   **RStudio**: *Ctrl+Alt+I (*or **Cmd + Option + I** on Mac)

## Demo: echo option

`echo` hides (FALSE)/shows (TRUE) code and outputs results

::::: columns
::: {.column width="45%"}
**Code**

```{{r}}
#| label: ex-orig
#| echo: true 
x = 1:10; y = 11:20
plot(x, y)
```
:::

::: {.column width="45%"}
**Output**

`ex-original` Output:

```{r, echo = TRUE}
#| label: ex-orig
#| fig-height: 2.5
x = 1:10; y = 11:20
plot(x, y)
```
:::
:::::

## Demo: echo option

::::: columns
::: {.column width="45%"}
**Code**

```{{r}}
#| label: ex-hide-code
#| echo: false
x = 1:10; y = 11:20
plot(x, y)
```
:::

::: {.column width="45%"}
**Output**

`ex-hide-code` Output:

```{r}
#| label: ex-hide-code
#| echo: false
#| fig-height: 2.5
x = 1:10; y = 11:20
plot(x, y)
```
:::
:::::

## Demo: eval option

`eval` controls whether the code is actually executed (`TRUE`) or just displayed without running (`FALSE`).

::::: columns
::: {.column width="45%"}
**Code**

```{{r}}
#| label: ex-orig
#| eval: true
x = 1:10; y = 11:20
plot(x, y)
```
:::

::: {.column width="45%"}
**Output**

`ex-original` Output:

```{r}
#| label: ex-orig-2
#| fig-height: 2.5
x = 1:10; y = 11:20
plot(x, y)
```
:::
:::::

## Demo: eval option

`eval` controls whether the code is actually executed (`TRUE`) or just displayed without running (`FALSE`).

::::: columns
::: {.column width="45%"}
**Code**

```{{r}}
#| label: ex-not-run
#| eval: false
x = 1:10; y = 11:20
plot(x, y)
```
:::

::: {.column width="45%"}
**Output**

`ex-not-run` Output:

```{r}
#| label: ex-not-run
#| eval: false
#| echo: true
x = 1:10; y = 11:20
plot(x, y)
```
:::
:::::

## Code Fold

```{r}
#| echo: true 
#| code-fold: true


x = 1:10; y = 11:20
plot(x, y)


```

# Code Chunk Exercise

# Inline Code {background-color="#E77500"}

## Inline code: what & why?

Insert results or (quick) calculations into running text, without additional text formatting. <br>

. . .

Why?

-   report and describe results reproducibly
-   minimizes errors: no manual copy-pasting results
-   automatic updating when data or code changes

## Inline code

Enclose the *R* expression using `` `r ` ``.

**Code:**

```{=html}
<div class="sourceCode">
<pre class="sourceCode markdown">
<code class="sourceCode markdown">
There are &#96;r nrow(cars)&#96; observations in our data. 
</code>
</pre>
</div>
```

**Output:**

::: border
There are `r nrow(cars)` observations in our data.
:::

::: callout-important
If using `Visual` or `Source` mode, be advised the *R* expression will only substitute the value held by the variable when the Quarto document is **rendered**. That is, the value contained within the expression only appears in the **output** file.
:::

## Reference code chunk variables inline

**Code:**

```{{{r}}}
#| label: calc-values
#| echo: false
x = 1:10
x_mu = mean(x)
x_sd = sd(x)
```

```{=html}
<div class="sourceCode">
<pre class="sourceCode markdown">
<code class="sourceCode markdown">
The _mean_ of **x** is  &#96;r x_mu&#96; and
the _standard deviation_ is &#96;r x_sd&#96;.
</code>
</pre>
</div>
```

**Output:**

::: border
The mean of **x** is 5.5 and the *standard deviation* is 3.02765.
:::

## Inline code: how?

**In R**

```{r}
#| label: Load the penguins dataset
#| echo: false
library(palmerpenguins)
```

``` md
There are `r knitr::inline_expr("sum(penguins$Species=='Adelie')")` 
penguins of the Adelie species in the dataset.
```

``` md
There are `r sum(penguins$species=="Adelie")` penguins of the "Adelie" species in the dataset.
```

# Inline Code Exercise
