---
title: "Workshop Setup"
subtitle: "Princeton Workshop: Reproducible Manuscripts in R with Quarto"
format:
  html: default
  docx: default
  pdf: default
bibliography: references.bib
editor: visual
---

## Before the workshop

Work through this document once, top to bottom. It installs everything we will use and then checks that it works.

1.  **Step 1:** Run each code chunk in the *Install* section by clicking the green ▶ (Run Current Chunk) button in its top-right corner. Wait for each one to finish (the red stop sign in the Console disappears) before running the next.
2.  **Step 2:** Click the **Render** button at the top of this editor. If an HTML page opens with a penguin plot, you are all set.

::: callout-note
The install chunks are marked `eval: false`, so they only run when *you* click ▶. Rendering will not reinstall anything.
:::

## Install

### R packages

This takes a few minutes the first time.

```{r}
#| label: install-packages
#| eval: false

install.packages(c(
  "tidyverse",      # data wrangling and plotting
  "palmerpenguins", # example dataset
  "quarto",         # run Quarto from R
  "rmarkdown",
  "knitr",
  "tinytex"         # LaTeX, for rendering to PDF
))
```

### LaTeX (for PDF output)

PDF output needs a LaTeX installation. TinyTeX is a small one that works well with Quarto. This also takes a few minutes.

```{r}
#| label: install-tinytex
#| eval: false

if (!tinytex::is_tinytex()) {
  tinytex::install_tinytex()
}
```

::: callout-tip
If you skip this step, everything except PDF output will still work.
:::

## Check

Everything below runs when you click **Render**.

```{r}
#| label: check-packages
#| message: false

pkgs <- c("tidyverse", "palmerpenguins", "quarto", "rmarkdown", "knitr", "tinytex")
installed <- pkgs %in% rownames(installed.packages())

data.frame(package = pkgs, installed = ifelse(installed, "yes", "MISSING"))
```

```{r}
#| label: check-tools
#| echo: false

quarto_version <- tryCatch(as.character(quarto::quarto_version()), error = function(e) "not found")
has_tinytex <- tinytex::is_tinytex()
```

-   R version: `r getRversion()`
-   Quarto version: `r quarto_version`
-   TinyTeX installed: `r if (has_tinytex) "yes" else "no (PDF output will not work yet)"`

### A plot

```{r}
#| label: fig-penguins
#| fig-cap: "Flipper length and body mass of Palmer penguins."
#| message: false
#| warning: false

library(tidyverse)
library(palmerpenguins)

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  labs(x = "Flipper length (mm)", y = "Body mass (g)", color = "Species") +
  theme_minimal()
```

### Inline code

The dataset has `r nrow(penguins)` penguins from `r n_distinct(penguins$island)` islands, with a mean body mass of `r round(mean(penguins$body_mass_g, na.rm = TRUE))` g (see @fig-penguins).

### A citation

The penguin data come from @gorman2014.

## Try other formats

Once the HTML version renders, try Word and PDF. Click the small arrow next to **Render** and pick **Render docx** or **Render pdf**.

## References
